---
title: "xkcd plot of german weather"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
#devtools::install_github('hadley/ggplot2')
library(plotly)
library(leaflet)
library(htmltools)
library(tidyverse)
library(xkcd)
library(extrafont)
library(ggrepel)
```

So [Maële Salmon](http://www.masalmon.eu/2017/11/16/wheretoliveus/) started this trend how to determine your city of choice based on your climate preference, inspired by [this xkcd graph](https://xkcd.com/1916/).

We already know that the [Spanish](https://twitter.com/claudiaguirao/status/931615734521909248) islands are our all-time favorite (you could have asked the German tourists) that you really should only go to [Iceland](https://twitter.com/matamix/status/932192147062784000) because of the landscape and that the weather is the same in the [Netherlands](https://twitter.com/RMHoge/status/932526164668829696), no matter where you go.

Different to the above mentioned analysis, I will just take a look at the average temperature to determine the ideal german city. The reason for this is the way the German Meteorological Service provides data. There is certainly a way to get the necessary data to calculate the [humidex](https://en.wikipedia.org/wiki/Humidex), but the time I want to put into this project is limited. For this reason, I reduce the analysis to the data set of average temperatures, downloaded from the [open data platform of the dwd](ftp://ftp-cdc.dwd.de/pub/CDC/regional_averages_DE/).

If you just forgot where Germany is located - here is a little help: 

```{r}
# Set coordinates (http://latitude.to/map/de/germany)
lng = 10.4541194
lat = 51.1642292

leaflet() %>% addTiles() %>%
  setView(lng=lng, lat=lat, zoom=4) %>%
  addProviderTiles(providers$Esri.NatGeoWorldMap) %>%
  addMarkers(lng=lng, lat=lat, 
             label = paste0("Coordinates: ",as.character(lat)," ",as.character(lng)))

#saveWidget(m, file="m.html")
```

## Import the Data

We have the seasonal average temperature as well as precipitation of the german federal states, derived from the gridded fields covering Germany for the period 1881 - 2017. 
```{r}
rm(list=ls())
air_temp <- read.csv("data/air_temperature_mean.txt", sep=";")
air_temp %>% gather(key = "location", value = "air_temp", -Jahr, -season) -> air_temp

rain <- read.csv("data/rain.txt", sep=";")
rain %>% gather(key = "location", value = "rain", -Jahr, -season) -> rain
```

## plot like xkcd 

I had to manually download the [xkcd font](http://simonsoftware.se/other/xkcd.ttf) and copy it to the font app on my Mac.

We use the fact that we have data from 1881 to the present day and compare 2016 with 1916:

```{r some data preperations}
seas <- c("summer", "winter")
delete <- c("Deutschland", "X")
air_temp %>% 
  filter(season %in% seas) %>%
  filter(!location %in% delete) %>%
  #filter(Jahr == 2016) %>%
  spread(season, air_temp) %>%
  filter(!is.na(winter)) %>%
  filter(!is.na(summer)) -> plot.df1
```

```{r scatterplot}
plot.df1 %>%
  filter(Jahr == 1916 | Jahr == 2016) %>%
  ggplot(aes(summer, winter, color=as.factor(Jahr))) +
  geom_point() +
  geom_text_repel(aes(label = location), family = "xkcd", 
                   max.iter = 50000, size = 3) +
  
  ggtitle("Where to live in Germany - 1916 vs. 2016",
          subtitle = "Data from DWD") +
  labs(color = "Year", 
       x= "Summer avg. temperature in Celsius degrees", 
       y = "Winter avg. temperature in Celsius degrees") +
  theme_xkcd() +
  theme(text = element_text(size = 13, family = "xkcd"))
```

### The entire course of time

Let's see if this average temperature rise can be tracked over time. Hover over the line to see data.

```{r}
seas <- c("summer", "winter")
delete <- c("Deutschland")

air_temp %>% 
  filter(season %in% seas) %>%
  filter(location %in% delete) -> plot.df2
```

```{r}
g <- plot.df2 %>%
  ggplot(aes(Jahr, air_temp, color=season)) +
  geom_line() +
  ggtitle("Average temperature in Germany - 1881-2017",
          subtitle = "Data from DWD") +
  labs(
       x= "", 
       y = "Average temperature in Celsius degrees") +
  theme_xkcd() +
  theme(text = element_text(size = 13, family = "xkcd"))

ggplotly(g, tooltip = c("Jahr","air_temp"))
```

